---
title: "HomUHet Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HomUHet Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

```
**Type** Package

**Title** Identifying and Separating Homogeneous and Heterogeneous Predictors

**Version**: 0.0.0.9000

**Date**: 2020-01-22

**Depends** tidyverse, glmnet, HDeconometrics, gglasso

**Suggests** knitr, rmarkdown

**Description** This package contains functions to identify and separate predictors with homogeneous or heterogeneous effects across ultra high-dimensional datasets.

**VignetteBuilder** knitr

**Encoding** UTF-8

**RoxygenNote** 7.1.1

**Author** Pei S. Yang [aut, cre]

**Maintainer** Pei S. Yang \<yang.1736@osu.edu\>

**Repository** Git Hub



### **R Topics Documented**



--------------------------------------
 *HomUHet*
--------------------------------------

**Description**

Identifies and separates predictors with homogeneous or heterogeneous effect across multiple data sets through fitting penalized linear regression models in two steps. Applicable to Gaussian response variable and very high dimensional data where the number of predictors could be larger than the number of observations per data set.

<br>
**Usage**

HomUHet (data, sid, solution_path=FALSE, y_name=NULL)



<br>
**Arguments**

**data** input matrix containing observations from all studies, with the Gaussian response variable vector as the first column, followed by the $N \times P$ predictor matrix with $P$ predictors and $N$ observations.

**sid** the study id for each observation in data. The sid argument needs to be a $N$ length vector containing natural numbers, where each entry will index the study from which the corresponding observation comes.

**solution_path** TRUE if the user wishes to output the solution path plots. Default is FALSE.

**y_name** if needed, supply a response variable name here to be displayed in the solution path plots. Default is NULL.

<br>
**Details**

Any further details such as standardization of variables, explanation of arguments or values should go here.

<br>
**Values**

**Homo** a vector containing the names or indexes of identified homogeneous predictors.

**Heter** a vector containing the names or indexes of identified heterogeneous predictors.

**Homo_estimates** a $p_{1}$ length vector containing the estimates of $p_{1}$ identified homogeneous predictors.

**Heter_estimates** a $p_{2} \times K$ matrix containing the estimates of $p_{2}$ identified heterogeneous predictors. 

**Solution_Path** solution path plots from the two steps.
       



